---
title: "Classification"
output: html_document
---

```{r}
data_path <- "Classification.csv"
cols <- c("Disability program", "Water resourses project", "Adoption project",	"Law on helthcare",	"Salvador project",	"Law on religions",
          "Antisatellite project",	"Nicaragua assistance project",	"Rocket project",	"Law on immigrants",	"Alternative fuel sources",	"Law on education",
          "Project on funds",	"Project on crime",	"Project of customs", 	"Project exports",	"Class")
senators <- read.csv(data_path, sep = '\t', header = FALSE, col.names = cols)
```


Split democrats and republicans:

```{r}
republicans <- senators[senators$Class == "republican",]
democrats <- senators[senators$Class == "democrat",]
```

Democrats outnumber republicans:

```{r}
barplot(table(senators$Class), col="grey")
```

There are more supporters for Law on religion than opponents
```{r}
barplot(table(senators$Law.on.religions), col="grey", main = "Law on religion")
```

```{r}
barplot(table(senators$Adoption.project), col="grey", main = "Adoption project")
```

```{r}
require(corrgram)

plotCorrelation <- function (data){
  
  data <- subset(data, select = -c(Class))
  
  data.reval <-  apply(data, 2, function(column){
    column <- as.numeric(factor(column, levels=c("no", "abstain", "yes")))
  })
  
  corrgram(data.reval, order = FALSE, cor.method = "kendall")
  #corrgram(cor.senator.reval, order = FALSE, cor.method = "kendall", lower.panel=panel.ellipse, upper.panel=panel.pie, text.panel=panel.txt)
}

```

We canHigh positive correlation between Nicaragua assistance and Adoption project. People who vote for them also voted for Rocket and Antisetellite projects. 
Another possible cluster may be formed by supporters for Law on healcare, Salvador project and Law on religion. This cluster is related to another possible group formed by 
supporters of Law on education, Project on funds and Project on crime. 

Two possible groups formed.

```{r}
plotCorrelation(senators)
```

Nicaragua and Rocket projects are mostly supported by democrats
```{r}
mosaicplot(senators$Class ~ senators$Nicaragua.assistance.project, shade=FALSE, color=TRUE, main="Nicaragua assistance project", xlab = "Senators", ylab ="Votes")
mosaicplot(senators$Class ~ senators$Rocket.project, shade=FALSE, color=TRUE, main="Rocket project", xlab = "Senators", ylab ="Votes")
```

Salvador and Funds mostly by republicans
```{r}
mosaicplot(senators$Class ~ senators$Salvador.project, shade=FALSE, color=TRUE, main="Salvador project", xlab = "Senators", ylab ="Votes")
mosaicplot(senators$Class ~ senators$Project.on.funds, shade=FALSE, color=TRUE, main="Project on funds", xlab = "Senators", ylab ="Votes")
```

Inside republicans two possible clusters can be seen: 
#1 Antisettellite proj, Nicaragua proj and Adoption project supporters
#2 Law on healhcare, Salvador, Law on religion
```{r}
plotCorrelation(republicans)
```

Democrats are more separeted. Supporters for Salvador/Law on religion have different opition with supporters Rocket and Nicaragua projects.
There is no one opition for those problems inside democrats. They look more separable than republicans
```{r}
plotCorrelation(democrats)
#add heat map for Salvador vs Rocket projects http://www.r-bloggers.com/search/heatmap
```

```{r}
clust.hierarchically <- function(data){
  library(cluster)
  library(stats)
  
  data.dist <- daisy(data, metric = "gower", stand = TRUE) 
  data.hc <- hclust(data.dist, method="complete")
  return(data.hc)
}
get.groups <- function(data, groups.num){
  library(stats)
  data.hc <- clust.hierarchically(data)
  groups <- cutree(data.hc, groups.num)
  return(cbind(data, groups))
}
```

```{r}
republicans.hc <- clust.hierarchically(republicans)
plot(republicans.hc, main = "Republicans dendrogram")
```

There is one senator #69 with totally different opinion. Event distant senators #71 and #149 vote for Antisatellite, Nicaragua, Salvador projects together.

```{r}
senators[c(69, 71, 149),]
```


```{r}
democrats.hc <- clust.hierarchically(democrats)
plot(democrats.hc, main = "Democrats dendrogram")
```

```{r}
senators[c(28, 54, 131),]
```

Group of republicans is very close to group of democrats
```{r}
senators.hc <- clust.hierarchically(subset(senators, select = -c(Class)))
senators.hc$labels <- senators$Class
#senators.hc$labels <- paste(as.character(senators.hc$order), as.character(senators$Class))
plot(senators.hc)
```

```{r}
get.number.of.clusters <- function(data) {
  library(fpc)
  library(cluster)
  
  data.dist <- daisy(data, metric = "gower", stand = TRUE)
  data.best <- pamk(data.dist)
  #cat("Number of clusters estimated by optimum average silhouette width:", data.best$nc, "\n")
  #visualized silhouette widthp
  #plot(pam(data.dist, data.best$nc))
  return(data.best)
}

clust.pam <- function(data, caption) {
  #par(mfrow=c(2,1))
  library(cluster)
  clust.num <- get.number.of.clusters(data)
  data.dist <- daisy(data, metric = "gower", stand = TRUE)
  pam.clust <- pam(data.dist, k=clust.num$nc)
  plot(pam(data.dist, clust.num$nc), main = paste("Silhouette plot for", caption))
  clusplot(pam.clust, main = paste("Partition around medoid for", caption)) #printed clusters
  return(pam.clust)
}
```

In general senators could be split in two groups

```{r}
senators.pam <- clust.pam(subset(senators, select = -c(Class)), 'senators')
```

Republicans more separable. Can be divedie in for clusters (with one obvious outlier)

```{r}
republicans.pam <- clust.pam(republicans, 'republicans')
```

Democrats are well devided in two groups

```{r}
democrats.pam <- clust.pam(democrats, 'democrats')
```





