---
title: "TimeSeries"
output: html_document
---

```{r, echo=FALSE}
library(forecast)
library(tseries)
```

```{r}
data_path <- "time-series.csv"
data <- read.csv(data_path, header = TRUE, sep = '\t')
```

```{r}
dat1_ts <- ts(data$Dat1)
```

```{r}
hist(data$Dat1, main = "Histogram for Data 1")
```

Looks like random data (white noise) with no trend

```{r}
plot(dat1_ts, main = "Data 1 time series")
```

No peaks in autocorrelation function (ACF), partial autocorrelation function (PACF) 
suggest that data does not depend on previous observations (stationary time series)

```{r}
tsdisplay(data$Dat1)
```

p-value for Dickey-Fuller suggest stationarity of the data

```{r}
adf.test(data$Dat1, alternative = c("stationary"))
```

KPSS Test

```{r}
kpss.test(data$Dat1)
```

Auto ARIMA selects ARIMA(0,0,0) model: 0 for ACF, PACF and 0 differences needed to make data stationary.

```{r}
dat1_mod_a <- auto.arima(dat1_ts)
summary(dat1_mod_a)
```

Forecast selects just expected value and adds margin confidence and prediction intervals 

```{r}
forecast_1a <- forecast(dat1_mod_a, 200)
plot(forecast_1a)
```

From observations to observation's growth for each next point (treating data as non stationary and trying to make it so for prediction)

```{r}
dat1_diff <- diff(data$Dat1)
```

Looks like previous value has effect in ACF and few previous values are valid for PACF (next value will be lower than previous)

```{r}
tsdisplay(dat1_diff)
```

But coefficients for args are not significant (division < 1)

```{r}
dat1_diff_ts <- ts(dat1_diff)
dat1_diff_mod_a <- auto.arima(dat1_diff_ts)
summary(dat1_diff_mod_a)
```

```{r}
hist(data$Dat2, main = "Histogram for Data 2")
```

Slowly decaying ACF suggests that data is not stationary. Clear positive trend

```{r}
#non-stationar data with trend, random walk
tsdisplay(data$Dat2)
dat2 <- data$Dat2
```

Stationarity hypothesis is rejected by Kwiatkowski-Phillips-Schmidt-Shin test

```{r}
dat2_ts <- ts(dat2)
adf.test(dat2, alternative = c("stationary"))
kpss.test(dat2)
```

First diff order to make data stationary

```{r}
dat2_diff <- diff(dat2)
```

Time series plot suggest stationarity. ACF and PACF show that value depends on previous values.

```{r}
tsdisplay(dat2_diff)
```

Auto ARIMA selects (1,1,2) model. Drift coef 1 suggests that diff was required.
All coef non zero - they are significant, drift 0.0138 suggests removed positive trend

```{r}
dat2_mod_a <- auto.arima(dat2_ts)
summary(dat2_mod_a)
```

Forecast for 100 values shows future positive trend

```{r}
forecast_2a <- forecast(dat2_mod_a, 100)
plot(forecast_2a)
```

Same for neural networks prediction

```{r}
nn_fit <- nnetar(dat2)
plot(forecast(nn_fit, 100))
```